---
title: "Experiment 3 - choice analysis (pre-registered dataset)"
author: 
  - name: "Zhang Chen"
    orcid: 0000-0002-3500-9182
date: "`r format(Sys.time(), '%d %B, %Y')`"
format: 
  html:
    code-fold: true
    code-tools: true
editor: visual
execute: 
  warning: false
  error: false
toc: true
toc-depth: 3
---

For this set of analyses, I use the pre-registered data set with 100 participants.

## Load libraries and data

```{r}
# load libraries
library(MASS)
library(Rmisc)
library(tidyverse)
library(ggpubr)

library(extraDistr)
library(loo)
library(bridgesampling)
library(brms)
library(cmdstanr)
library(bayesplot)
library(bayestestR)
library(tidybayes)
library(sjPlot)

# parallelize the chains using all the cores
options(mc.cores = parallel::detectCores())

# set the theme for all ggplot2 figures
theme_set(theme_bw() +
            theme(legend.position = "top",
                  legend.direction = "horizontal"))

# create a folder for saving brms fits
if(!dir.exists("brms-fits")){
  dir.create("brms-fits")
}

# load the cleaned data
df_main <- read_csv("../../../data/processed/df_main_exp3_prereg.csv")
```

## Choices

### Catch trials

In this experiment, I included two types of catch trials, either the high-probability option or the low-probability option has a much higher expected value than the other (in the latter LP-optimal catch trials, the HP option actually has an EV of 0). Here I plot how often participants chose the higher EV option for these two types of catch trials (@fig-catch-choice).

```{r}
#| label: fig-catch-choice
#| fig-cap: The proportion of choosing the high EV option on the catch trials when the high-probability option (top) or the low-probability option (bottom) has a higher EV.
#| fig-width: 7
#| fig-height: 5 

# catch trials
df_catch_choices <- df_main %>%
  filter(trial_type == "catch") %>%
  group_by(subject_ID, high_EV_option) %>%
  summarize(
    # the total number of trials, should be 12 for all participants
    n_total = n(),
    # the number of trials on which people chose the high EV option
    n_high_EV =  sum(choose_high_EV == "yes"),
    # the proportion of choosing the high EV option
    prop_high_EV = n_high_EV / n_total* 100
  ) %>%
  ungroup()

# plot the choices on the catch trials
ggplot(df_catch_choices, aes(subject_ID, prop_high_EV)) +
  geom_line(aes(group = 1), color = "gray") +
  geom_point() +
  labs(x = "Participant", 
       y = "The proportion of choosing high EV options") +
  facet_wrap(~high_EV_option, ncol = 1)

# compute some descriptives
catch_choices_descriptive <- df_catch_choices %>%
  group_by(high_EV_option) %>%
  summarize(
    mean = mean(n_high_EV),
    sd = sd(n_high_EV),
    min = min(n_high_EV),
    max = max(n_high_EV)
  )
```

### Experimental trials

#### EV ratios

In this analysis, I use the EV ratio between two options, the outcome of game 1 (loss = 0.5, win = -0.5) and whether there was a pause or not (yes = 0.5, no = -0.5) and their interactions as predictors to predict whether participants chose the HP option or not on each trial. I use the maximum random structure by including both random intercept and all random slopes per participant.

```{r}
# find the participants that need to be excluded,
# here I do not exclude any participants
high_prob_cutoff <- 9
low_prob_cutoff <- 9

subject_exclude1 <- df_catch_choices %>%
  filter(
    high_EV_option == "HP",
    n_high_EV < high_prob_cutoff
    ) %>% .$subject_ID

subject_exclude2 <- df_catch_choices %>%
  filter(
    high_EV_option == "LP",
    n_high_EV < low_prob_cutoff
    ) %>% .$subject_ID

df_exp_selected <- df_main %>%
  filter(
    !subject_ID %in% subject_exclude1,
    !subject_ID %in% subject_exclude2,
    trial_type == "exp"
  )

# fit a brms model - go for the maximum random structure
brms_file_name <- 
  paste("brms-fits/fit_exp_HP", 
        high_prob_cutoff, "LP", low_prob_cutoff, sep = "_")

fit_exp <- brm(
  game2_choose_HP ~ EV_ratio * game1_outcome_num * delay_num +
    (EV_ratio * game1_outcome_num * delay_num|subject_ID),
  family = bernoulli(link = "logit"),
  prior = 
    c(
      prior(normal(0, 2), class = Intercept),
      prior(normal(0, 1), class = b),
      prior(normal(0, 1), class = sd),
      prior(lkj(2), class = cor)
    ),
  data = df_exp_selected,
  cores = 4,
  iter = 15000,
  warmup = 5000,
  seed = 1234,
  file = brms_file_name,
  backend = "cmdstanr"
)

# save the data for later analysis
write_csv(df_exp_selected, 
          "../../../data/processed/df_exp_exp3_prereg.csv")

```

Some general posterior predictive checks (@fig-ppc-general).

```{r}
#| label: fig-ppc-general
#| fig-cap: General posterior predictive checks.
#| fig-subcap: 
#| - "As a function of delay"
#| - "As a function of game 1 outcome"
#| - "As a function of EV ratio"
#| layout: [[1, 1], [1]]
#| fig-width: 8
#| fig-height: 8


# model diagnostics
# check the trace plot, looks good.
# plot(fit_exp)

# posterior predictive checks
yrep <- posterior_predict(fit_exp)

ppc_bars_grouped(df_exp_selected$game2_choose_HP, yrep, 
                 df_exp_selected$delay, prob = 0.95, freq = FALSE)

ppc_bars_grouped(df_exp_selected$game2_choose_HP, yrep, 
                 df_exp_selected$game1_outcome, prob = 0.95, freq = FALSE)

ppc_bars_grouped(df_exp_selected$game2_choose_HP, yrep, 
                 df_exp_selected$EV_ratio, prob = 0.95, freq = FALSE)
```

Make a plot to show the effects of EV ratio and delay.

```{r}
#| label: fig-EV-ratio-delay
#| fig-cap: Probability of choosing the HP option as a function of EV ratio and pause.
#| fig-width: 6
#| fig-height: 6

# Here I will plot the probability of choosing the HP option
# as a function of the EV ratio and whether there was a delay or not
# EV ratio will be the x axis, and delay (yes vs. no) as two colors
# here I define a small adjustment value to the x axis (i.e. EV ratio),
# that moves all no delay condition data points a bit to the left,
# and the delay condition data points to the right, to reduce overlapping.
x_adj_amount <- 0.025

# compute the probability of choosing HP for each participant
# in each condition.
plot_raw_data <- df_exp_selected %>%
  group_by(subject_ID, EV_ratio, delay) %>%
  summarize(p_HP = mean(game2_choose_HP)) %>%
  mutate(
    # add a bit adjustment as mentioned above
    EV_ratio_adj = ifelse(delay == "yes", x_adj_amount, -x_adj_amount),
    EV_ratio_plot = EV_ratio + EV_ratio_adj
  )

# get the predicted prob of choosing HP from brms
int_conds <- list(
  EV_ratio = unique(df_exp_selected$EV_ratio),
  delay_num = c(-0.5, 0.5)
)

plot_prediction <- 
  conditional_effects(
    fit_exp, 
    "EV_ratio:delay_num",
    int_conditions = int_conds)

# turn the predictions into a tibble
plot_pred <- as_tibble(
  plot_prediction$`EV_ratio:delay_num`
  ) %>%
  mutate(
    # similarly, add a little adjustment to the EV ratio
    delay = ifelse(delay_num == 0.5, "yes", "no"),
    EV_ratio_adj = ifelse(delay == "yes", x_adj_amount, -x_adj_amount),
    EV_ratio_plot = EV_ratio + EV_ratio_adj
  )

# plot the results
EV_delay_plot <- plot_raw_data %>%
  ggplot(aes(EV_ratio_plot, p_HP, color = delay)) +
  # plot raw data per participant as transparent open circles
  # jitter the dots vertically a bit to reduce overlapping
  geom_jitter(alpha = 0.2, height = 0.05, shape = 1) +
  # connect the predicted prob of choosing HP with lines
  geom_line(data = plot_pred, 
            aes(x = EV_ratio_plot,
                y = estimate__,
                group = delay),
            alpha = 0.4) +
  # estimates as solid points
  geom_point(data = plot_pred, 
             aes(EV_ratio_plot, 
                 estimate__)) +
  # add 95% CI as error bars
  geom_errorbar(data = plot_pred,
                aes(x = EV_ratio_plot, 
                    y = estimate__,
                    ymin = lower__, 
                    ymax = upper__),
                width = 0.05) +
  labs(x = "EV ratio", 
       y = "P(Choose the HP option)",
       color = "Pause")

EV_delay_plot

# save as a png file
ggsave(filename = "../5-plots/EV_delay_Exp3_reg.png", 
       EV_delay_plot,
       width = 4, height = 4)
```

Make a similar plot, but this time also take the previous outcome into account.

```{r}
#| label: fig-EV-ratio-delay-outcome
#| fig-cap: Probability of choosing the HP option as a function of EV ratio, pause, and previous outcome.
#| fig-width: 7
#| fig-height: 5

source("function/choice_plot_functions.R")

# select parameters of interest
var_selected <- get_variables(fit_exp)[1:8]

# get posterior draws for these parameters
draws <- fit_exp %>% as_draws_df(variable = var_selected) 

choice_params_plot <- plot_choice_params(draws)

choice_plots <- plot_choice(draws)

choice_pause_outcome_plot <- choice_plots$p1

choice_pause_outcome_plot

ggsave("../plots/choice_EV_exp3_prereg.png",
       choice_pause_outcome_plot,
       width = 5.5, height = 3.5)
```

```{r, results='asis'}
tab_model(fit_exp)
```

#### Prob and amount

```{r}
# compute the prob ratio and amount ratio between two options
df_exp_selected <- df_exp_selected %>%
  mutate(
    prob_ratio = (game2_HP_prob - game2_LP_prob)/(game2_HP_prob + game2_LP_prob) * 2,
    amount_ratio = (game2_HP_amount - game2_LP_amount)/(game2_HP_amount + game2_LP_amount) * 2
  )

# get the unique trials and compute the mean prob_ratio and amount_ratio
df_unique <- df_exp_selected %>%
  select(game2_HP_prob, game2_HP_amount, game2_LP_prob, game2_LP_amount, prob_ratio, amount_ratio) %>%
  unique()

prob_ratio_mean <- mean(df_unique$prob_ratio)
amount_ratio_mean <- mean(df_unique$amount_ratio)

# center the prob_ratio and amount_ratio predictors
df_exp_selected <- df_exp_selected %>%
  mutate(
    prob_ratio_centered = prob_ratio - prob_ratio_mean,
    amount_ratio_centered = amount_ratio - amount_ratio_mean
  )

# fit a brms model
brms_file_name <- "brms-fits/fit_prob_amount"

fit_prob_amount <- brm(
  game2_choose_HP ~ prob_ratio_centered * amount_ratio_centered *
                    game1_outcome_num * delay_num +
                    (prob_ratio_centered * amount_ratio_centered *
                    game1_outcome_num * delay_num|subject_ID),
  family = bernoulli(link = "logit"),
  prior = 
    c(
      prior(normal(0, 2), class = Intercept),
      prior(normal(0, 1), class = b),
      prior(normal(0, 1), class = sd),
      prior(lkj(2), class = cor)
    ),
  data = df_exp_selected,
  cores = 4,
  iter = 15000,
  warmup = 5000,
  seed = 1234,
  file = brms_file_name,
  backend = "cmdstanr"
)
```

```{r}
#| label: fig-ppc-general-2
#| fig-cap: General posterior predictive checks.
#| fig-subcap: 
#| - "As a function of delay"
#| - "As a function of game 1 outcome"
#| - "As a function of EV ratio"
#| layout: [[1, 1], [1]]
#| fig-width: 8
#| fig-height: 8

# model diagnostics
# check the trace plot, looks good.
# plot(fit_prob_amount)

# posterior predictive checks
yrep <- posterior_predict(fit_prob_amount)

ppc_bars_grouped(df_exp_selected$game2_choose_HP, yrep, 
                 df_exp_selected$delay, prob = 0.95, freq = FALSE)

ppc_bars_grouped(df_exp_selected$game2_choose_HP, yrep, 
                 df_exp_selected$game1_outcome, prob = 0.95, freq = FALSE)

ppc_bars_grouped(df_exp_selected$game2_choose_HP, yrep, 
                 df_exp_selected$EV_ratio, prob = 0.95, freq = FALSE)
```

```{r}
var_selected <- get_variables(fit_prob_amount)[1:16]

draws <- fit_prob_amount %>%
  as_draws_df(variable = var_selected) %>%
  pivot_longer(cols = all_of(var_selected),
               names_to = "parameter",
               values_to = "estimate") %>%
  mutate(parameter = 
           factor(parameter, levels = var_selected,
                  labels = c("Intercept",
                             "Probability",
                             "Amount",
                             "Prev outcome (loss vs. win)",
                             "Pause (yes vs. no)",
                             "Prob * Amount",
                             "Prob * Prev outcome",
                             "Amount * Prev outcome",
                             "Prob * Pause",
                             "Amount * Pause",
                             "Prev outcome * Pause",
                             "Prob * Amount * Prev outcome",
                             "Prob * Amount * Pause",
                             "Prob * Prev outcome * Pause",
                             "Amount * Prev outcome * Pause",
                             "Prob * Amount * Prev outcome * Pause"))
  )

prob_amount_post_plot <- draws %>%
  ggplot(aes(x = estimate, y = parameter)) +
  stat_halfeye() +
  geom_vline(xintercept = 0, linetype = "dashed") +
  labs(x = "Posterior distribution (log odds-ratio)", y = "")+
  scale_y_discrete(limits=rev) +
  theme(axis.text = element_text(size = 10, color = "black"),
        axis.title = element_text(size = 11))

prob_amount_post_plot

# another version with only a few variables of interest
param_selected <- c("Probability", "Prob * Prev outcome",
                    "Prob * Pause", "Amount",
                    "Amount * Prev outcome", "Amount * Pause")

prob_amount_post_selected_plot <- draws %>%
  filter(parameter %in% param_selected) %>%
  mutate(parameter = factor(parameter, levels = param_selected)) %>%
  ggplot(aes(x = estimate, y = parameter)) +
  stat_halfeye() +
  geom_vline(xintercept = 0, linetype = "dashed") +
  labs(x = "Posterior distribution (log odds-ratio)", y = "")+
  scale_y_discrete(limits=rev) +
  theme(axis.text = element_text(size = 10, color = "black"),
        axis.title = element_text(size = 11))

# save as a png file
ggsave(filename = "../5-plots/prob_amount_Exp3_reg.png", 
       prob_amount_post_plot,
       width = 7, height = 8)
```

```{r, results='asis'}
tab_model(fit_prob_amount)
```

Compare this model with probability and amount ratios as two separate predictors, with the model from above with only the EV ratio as the key predictor.

```{r, cache=TRUE}
# compute the loo model fit criteria

# loo_exp <- loo(fit_exp, pointwise = TRUE)
# loo_prob_amount <- loo(fit_prob_amount, pointwise = TRUE)

# loo_compare(fit_exp, fit_prob_amount, criterion = "loo")
```

## Between-exp comparisons

### With Experiment 2

Compare the current experiment with Experiment 2.

```{r}
# load the data from Experiment 2
df_exp2 <- read_csv("../../../../Exp2/data/processed/df_exp_exp2.csv")

# add experiment number, and adjust the subject ID
df_exp2 <- df_exp2 %>%
  mutate(
    Exp = "Exp2",
    subject_ID = paste0(Exp, subject_ID, sep = "_")
  )

# load the data from the current experiment again
df_exp3 <- read_csv("../../../data/processed/df_exp_exp3_reg.csv")

df_exp3 <- df_exp3 %>%
  mutate(
    Exp = "Exp3",
    subject_ID = paste0(Exp, subject_ID, sep = "_")
  )

# combine data from both experiments
df_exp2_exp3 <- df_exp2 %>%
  bind_rows(df_exp3) %>%
  mutate(Exp_num = ifelse(Exp == "Exp2", 0.5, -0.5))
```

```{r}
# fit a brms model
brms_file_name <- "brms-fits/fit_choice_exp2_exp3"

fit_choice_exp2_exp3 <- brm(
  game2_choose_HP ~ EV_ratio * game1_outcome_num * delay_num * Exp_num +
    (EV_ratio * game1_outcome_num * delay_num|subject_ID),
  family = bernoulli(link = "logit"),
  prior = 
    c(
      prior(normal(0, 2), class = Intercept),
      prior(normal(0, 1), class = b),
      prior(normal(0, 1), class = sd),
      prior(lkj(2), class = cor)
    ),
  data = df_exp2_exp3,
  cores = 4,
  iter = 15000,
  warmup = 5000,
  seed = 1234,
  file = brms_file_name,
  backend = "cmdstanr"
)
```

Again compare the current experiment with Experiment 2, but this time using probability and amount ratio as predictors.

```{r}
# compute the prob ratio and amount ratio between two options
df_exp2_exp3 <- df_exp2_exp3 %>%
  mutate(
    prob_ratio = (game2_HP_prob - game2_LP_prob)/(game2_HP_prob + game2_LP_prob) * 2,
    amount_ratio = (game2_HP_amount - game2_LP_amount)/(game2_HP_amount + game2_LP_amount) * 2
  )

# get the unique trials and compute the mean prob_ratio and amount_ratio
df_unique <- df_exp2_exp3 %>%
  select(game2_HP_prob, game2_HP_amount, game2_LP_prob, game2_LP_amount, prob_ratio, amount_ratio) %>%
  unique()

prob_ratio_mean <- mean(df_unique$prob_ratio)
amount_ratio_mean <- mean(df_unique$amount_ratio)

# center the prob_ratio and amount_ratio predictors
df_exp2_exp3 <- df_exp2_exp3 %>%
  mutate(
    prob_ratio_centered = prob_ratio - prob_ratio_mean,
    amount_ratio_centered = amount_ratio - amount_ratio_mean
  )

# fit a brms model
brms_file_name <- "brms-fits/fit_prob_amount_exp2_exp3"

fit_prob_amount_exp2_exp3 <- brm(
  game2_choose_HP ~ prob_ratio_centered * amount_ratio_centered *
                    game1_outcome_num * delay_num * Exp_num +
                    (prob_ratio_centered * amount_ratio_centered *
                    game1_outcome_num * delay_num|subject_ID),
  family = bernoulli(link = "logit"),
  prior = 
    c(
      prior(normal(0, 2), class = Intercept),
      prior(normal(0, 1), class = b),
      prior(normal(0, 1), class = sd),
      prior(lkj(2), class = cor)
    ),
  data = df_exp2_exp3,
  cores = 4,
  iter = 15000,
  warmup = 5000,
  seed = 1234,
  file = brms_file_name,
  backend = "cmdstanr"
)
```

### With Experiment 1

```{r}
# load the data from Experiment 1
df_exp1 <- read_csv("../../../../Exp1/data/processed/df_exp_exp1.csv")

# add experiment number, and adjust the subject ID
df_exp1 <- df_exp1 %>%
  mutate(
    Exp = "Exp1",
    subject_ID = paste0(Exp, subject_ID, sep = "_")
  )

# load the data from the current experiment again
df_exp3 <- read_csv("../../../data/processed/df_exp_exp3_reg.csv")

df_exp3 <- df_exp3 %>%
  mutate(
    Exp = "Exp3",
    subject_ID = paste0(Exp, subject_ID, sep = "_")
  )

# combine data from both experiments
df_exp1_exp3 <- df_exp1 %>%
  bind_rows(df_exp3) %>%
  mutate(Exp_num = ifelse(Exp == "Exp1", 0.5, -0.5))
```

```{r}
# fit a brms model
brms_file_name <- "brms-fits/fit_choice_exp1_exp3"

fit_choice_exp1_exp3 <- brm(
  game2_choose_HP ~ EV_ratio * game1_outcome_num * delay_num * Exp_num +
    (EV_ratio * game1_outcome_num * delay_num|subject_ID),
  family = bernoulli(link = "logit"),
  prior = 
    c(
      prior(normal(0, 2), class = Intercept),
      prior(normal(0, 1), class = b),
      prior(normal(0, 1), class = sd),
      prior(lkj(2), class = cor)
    ),
  data = df_exp1_exp3,
  cores = 4,
  iter = 15000,
  warmup = 5000,
  seed = 1234,
  file = brms_file_name,
  backend = "cmdstanr"
)
```

Compare the current experiment with Experiment 1, with probability and amount ratios as predictors.

```{r}
# compute the prob ratio and amount ratio between two options
df_exp1_exp3 <- df_exp1_exp3 %>%
  mutate(
    prob_ratio = (game2_HP_prob - game2_LP_prob)/(game2_HP_prob + game2_LP_prob) * 2,
    amount_ratio = (game2_HP_amount - game2_LP_amount)/(game2_HP_amount + game2_LP_amount) * 2,
    prob_ratio_centered = prob_ratio - prob_ratio_mean,
    amount_ratio_centered = amount_ratio - amount_ratio_mean
  )

# fit a brms model
brms_file_name <- "brms-fits/fit_prob_amount_exp1_exp3"

fit_prob_amount_exp1_exp3 <- brm(
  game2_choose_HP ~ prob_ratio_centered * amount_ratio_centered *
                    game1_outcome_num * delay_num * Exp_num +
                    (prob_ratio_centered * amount_ratio_centered *
                    game1_outcome_num * delay_num|subject_ID),
  family = bernoulli(link = "logit"),
  prior = 
    c(
      prior(normal(0, 2), class = Intercept),
      prior(normal(0, 1), class = b),
      prior(normal(0, 1), class = sd),
      prior(lkj(2), class = cor)
    ),
  data = df_exp1_exp3,
  cores = 4,
  iter = 15000,
  warmup = 5000,
  seed = 1234,
  file = brms_file_name,
  backend = "cmdstanr"
)
```
